Analyzing Influencing Factors on Family Size in Portugal

### 2025-02-11 paper/paper.rmd
Completed

The Poisson GLM was developed to investigate the relationship between literacy and age at marriage while accounting for the natural log of the duration of marriage through an offset term. The model includes literacy status and categorized age at marriage as predictor variables, with the reference group consisting of literate individuals who married between the ages of 22 and 25. The results (Table \@ref(tab:model-p)) indicate that illiteracy is a significant predictor (p = 0.000), suggesting that being illiterate is strongly positively associated with changes in the family size compared to literate individuals. Among the marriage age groups, individuals who married between 15-18 years (p = 0.025) and 18-20 years (p = 0.026) show significant differences from the reference category, implying that marrying at these ages has a notable impact. Conversely, marriage before 15, between 20-22, between 25-30, and after 30 do not show significant differences from the reference group. These findings highlight the role of literacy and specific marriage age ranges in influencing the outcome, though additional modeling techniques or interactions could be explored to refine the analysis further.

@@

>>>>

The Poisson GLM was developed to investigate the relationship between literacy and age at marriage while accounting for the natural log of the duration of marriage through an offset term. The model includes literacy status and categorized age at marriage as predictor variables, with the reference group consisting of literate individuals who married between the ages of 22 and 25. The results (Table \@ref(tab:model-p)) indicate that illiteracy is a significant predictor (p = 0.000), suggesting that being illiterate is strongly positively associated with changes in the family size compared to literate individuals. Among the marriage age groups, individuals who married between 15-18 years (p = 0.025) and 18-20 years (p = 0.026) show significant differences from the reference category, implying that marrying at these ages has a notable impact. Conversely, marriage before 15, between 20-22, between 25-30, and after 30 do not show significant differences from the reference group. These findings highlight the role of literacy and specific marriage age ranges in influencing the outcome, though additional modeling techniques or interactions could be explored to refine the analysis further. Also, effects of the predictors on family size could be seen clearly in Figure \@ref(fig:effect).

@@

```{r effect, fig.cap="Comparison of the relative risk (RR) estimates and 95% confidence intervals from a Poisson model examining literacy status and age at marriage. The reference category for literacy is literate, while different age groups at marriage are compared against a baseline. The results indicate no significant differences in RR across age groups, as all confidence intervals include 1. The effect of literacy shows a slightly larger estimate but remains within the range of non-significance at the 0.05 significance level.", fig.pos="H", fig.height=6, fig.width=8, message=FALSE, echo=FALSE}
poi_modelCI <-  as.data.frame(exp(confint(model_p)[-1,]))
poi_modelCI$level <- gsub(paste(names(model_p$xlevels), collapse='|'),
   "", rownames(poi_modelCI))
poi_modelCI$variable=unlist(strsplit(rownames(poi_modelCI),poi_modelCI$level))
poi_modelCI$x <- 1:nrow(poi_modelCI)
poi_modelCI$cex <- sqrt(1/apply(log(poi_modelCI[,1:2]),1,diff))
forXaxis = tapply(poi_modelCI$x, poi_modelCI$variable, mean)
matplot(poi_modelCI[,1:2], type='n', xaxt='n', bty='n', xaxt='n', log='y', ylab='RR')
segments(poi_modelCI$x, poi_modelCI[,1],
   poi_modelCI$x, poi_modelCI[,2])
numeric_coef <- as.numeric(as.character(poi_modelCI$coef[-1]))
numeric_coef <- rep(0, length(poi_modelCI$x))
 points(poi_modelCI$x,
   exp(numeric_coef), pch=15,
   cex = poi_modelCI$cex)

mtext(poi_modelCI$level, 1, at=poi_modelCI$x,
   las=3, line=-1)
mtext(names(forXaxis), 1, at=forXaxis,
   line=-2)
abline(h=1, lty=3)
mtext(poi_modelCI$level, side = 1, at = poi_modelCI$x, las = 1, line = 3, cex = 0.5)
```
